---
title: "ExerciseData"
date: "19/01/2022"
output: html_document
---
We will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.
```{r}
library(lattice)
library(ggplot2)
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)

set.seed(1234)

trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"

testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

trainingset <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testingset <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

#trainingset <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
#testingset <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))

#dim(trainingset); dim(testingset); summary(trainingset); summary(testingset); str(trainingset); str(testingset); head(trainingset); head(testingset);               

trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]

trainingset   <-trainingset[,-c(1:7)]
testingset <-testingset[,-c(1:7)]

traintrainset <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
TrainTrainingSet <- trainingset[traintrainset, ] 
TestTrainingSet <- trainingset[-traintrainset, ]

plot(TrainTrainingSet$classe, col="yellow", main="Plot of levels of variable classe within the TrainTrainingSet data set", xlab="classe", ylab="Frequency", ylim=c(0,300))

```

Above code shows that each level frequency is within the same order of magnitude of each other. Level A is the most frequent while level D is the least frequent.

```{r}
model1 <- rpart(classe ~ ., data=TrainTrainingSet, method="class")

prediction1 <- predict(model1, TestTrainingSet, type = "class")

rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)

confusionMatrix(
  factor(prediction1, levels = 1:148),
  factor(TestTrainingSet$classe, levels = 1:148)
)

TrainTrainingSet$classe = factor(TrainTrainingSet$classe)
model2 <- randomForest(classe ~. , data=TrainTrainingSet, method="class")

prediction2 <- predict(model2, TestTrainingSet, type = "class")

confusionMatrix(prediction2, TestTrainingSet$classe)

#final result
predictfinal <- predict(model2, testingset, type="class")
predictfinal
```

Which prediction model to use?
Random Forest algorithm performed better than Decision Trees. Accuracy for Random Forest model was 0.995 (95% CI: (0.993, 0.997)) compared to Decision Tree model with 0.739 (95% CI: (0.727, 0.752)). The Random Forests model is choosen. The expected out-of-sample error is estimated at 0.005, or 0.5%.
